---
title: "Statistical Documentation with R Markdown"
author: "Katherine Simeon"
date: "May 29, 2018"
output: 
  ioslides_presentation:
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# Make sure packages are installed
library(languageR) 
library(lme4)
library(ggplot2)
library(car)
data("lexdec") # This is the data we are using today!
```

## Why is statistical documentation important? 

>- Reproducibility  
>- Sanity of future self
>- Good communication practice

## Challenges when documenting your work  

>- Missing information (assumes you'll remember something but you actually won't)       
>- Many moving parts        
>- Time consuming (to create and maintain)        

## What makes good documentation? {.build}
  
Varies/depends!  
  
>- Who's the intended **audience**?  (future you vs. fellow labmate vs. external person)     
>- What's the intended **use**? (how specific or generalizable should it be?)       
>- Statistical Software (in this example, I'll be using R)      
      
Regardless, there should still be a generally standardized format:  

>- Version control / revision history            
>- Software dependencies and what files are needed (and what format they should be in)     

## Documentation components  
  
>- Data & the process for cleaning it    
>- Description of IVs & DVs (how they're quantified, levels/conditions for each variable, etc.)     
>- Syntax for statistical analyses    
>- Syntax for figures (if applicable)  
>- Thorough commenting

## The lowdown on R Markdown 
  
- Creates reproducible, dynamic documents with R code embedded  
- Technically markdown, but specific to **R**    

## Getting started with R Markdown     
  
```{r, echo=TRUE, eval=FALSE}
install.packages('rmarkdown')
```
  
>1. Installing package should automatically install dependencies       
>2. In drop-down menu, create new .Rmd file        
>3. Choose output format (this can also be modified later in the header)     
>4. Create your document (this [cheatsheet](https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf) is a great place to start)     
>5. "Knit" it into desired output format!      

## R Markdown Basics {.build} 
  
R code is preceded by: ```{r, *a bunch of arguments*}```    
     
_The important arguments are:_    
  
```echo=TRUE or FALSE```  
Should the code be _included_ in the final document?   
  
```eval=TRUE or FALSE```  
Should the code be _evaluated_ in the final document?           
  
## Setting up your document {.build}  
  
Every new file has a ```{r setup}```, where you should:
  
- Set up your working directory   
- Load necessary packages     
- This chunk is run automatically when you run any other chunk in your document! (so this is a good place to load your data in too)  
  
  
My set up for this presentation looks like this:  
```{r, echo=TRUE, eval=FALSE}
# Make sure packages are installed
library(languageR) 
library(lme4)
library(ggplot2)
library(car)
data("lexdec") # This is the data we are using today!
```

## Our data {.build}  
  
_**lexdec**_ - Lexical decision latencies elicited from 21 subjects for 79 English concrete nouns.  
   
- _Native Language_ is classified for each subject as **English** or **Other**.     
- Each word is sorted into two levels of _semantic class_: **animal** or **plant**.
  
From [languageR package](https://cran.r-project.org/web/packages/languageR/languageR.pdf)  

## lexdec {.build}
  
```{r, echo=TRUE, eval=TRUE}
# This dataset has a lot of columns
dim(lexdec)

# Only printing the first 6 rows of the first 8 columns 
head(lexdec[,1:8])
```

## Example: A mixed effects model {.build} 
  
```{r, echo=FALSE, eval=TRUE}
# Setting up contrasts and centering data for analysis  
# Set contrasts & types for variables of interest: semantic class ('animal' / 'plant') and native language ('english' / 'other'). 
contrasts(lexdec$Class) <- c(.5,-.5)
contrasts(lexdec$NativeLanguage) <- c(.5,-.5)
#'center' the continuous variable we're using so that the mean is at 0
lexdec$FrequencyC <- lexdec$Frequency - mean(lexdec$Frequency)
# also calculate numeric versions of Class and NativeLanguage that have the values of .5 and -.5
lexdec$ClassN <- (as.numeric(lexdec$Class)-1.5)
lexdec$NativeLanguageN <- (as.numeric(lexdec$NativeLanguage)-1.5)
```

In this model, we want to predict reaction time (RT) from the fixed effects of native language, and semantic class.  
  
We want to account for random effects of subjects and items (words).    

```{r, echo=TRUE, eval=TRUE}
model <- lmer(RT ~ NativeLanguageN*ClassN # fixed effects on DV
              + (1+ClassN|Subject) # Random effects of subject 
              + (1+NativeLanguageN|Word), # Random effects of item
              data=lexdec,REML=F)  
```
  
_Thank you to Laurel Brehm for teaching us this 2 years ago!_  
_See her tutorial [here](https://sites.psu.edu/laurelbrehm/scripts-stats/)_

## Sorry output does not fit on the slide  

(And I was a too lazy to adjust it...)  
```{r, echo=TRUE, eval=TRUE}
summary(model) # Let's see the model
```
  
----  
```{r, echo=TRUE, eval=TRUE}
Anova(model) # Let's see p-values
```

## Let's make a figure! {.build}
  
Let's plot RT based on word frequency, sorted by native language and semantic class.  
  
```{r, echo=FALSE, eval=TRUE}
ggplot(aes(x = Frequency, y = RT, colour = NativeLanguage),
       data = lexdec) +
  geom_point() +
  scale_x_continuous("Written log frequency (CELEX)") +
  scale_y_continuous("Lexical decision log RTs") +
  facet_grid(.~Class)
```

## Advantages of documenting in R Markdown  
  
- Accomodates large chunks of text (no rows/paragraphs of #)       
- Can organize your code into runnable chunks     
- Can be "knit" into multiple formats (pdf, MS Word, HTML)       

## Other fun R things (that are helpful for documentation)  {.build}
  
**R Projects**        

- Each project is in its own directory & workspace          
- Easy to share multiple files with collaborators     
  
**roxygen skeletons**         

- Provides the template that is used to document functions        
- *Code* > *Insert Roxygen Skeleton*    
- Package: roxygen2     
